ニューラルネットを用いた逆コンパイラとその評価

### 2020-05-26 texs/paper_base.tex
add broken pdf

\documentclass[senior,final,11pt]{iscs-thesis}
%論文の種類とフォントサイズをオプションに
%\usepackage{graphicx}% 必要に応じて
%\usepackage{mysettings}% 自分用設定

@@

%-------------------
\etitle{Neural Network based Decompiler and its Evaluation}
\jtitle{ニューラルネットを用いた逆コンパイラとその評価}

__NAMES__

@@

\usepackage{listings}
\usepackage{cite}
\usepackage{url}
\usepackage[numbers]{natbib}
\usepackage[dvipdfmx]{graphicx}
\usepackage{framed}
\usepackage{caption}

@@

\begin{document}
\begin{eabstract}
A decompiler is a tool for recovering a source code from compiled binary data.
There are various decompilers, but they often output a source code that is not intelligible to humans. 
In this thesis, we apply machine translation techniques to generate human intelligible decompiled source codes. 
More specifically, we propose to use recurrent neural networks with attention, 
which have been demonstrated to be useful for machine translation. 
In experiments, we use source codes collected from open source projects and their binary data for training and evaluating the neural networks.
From the experiment, we can observe little improvement in the result of decompilation with the use of attention.
Unfortunately, the results were not still good enough for practical use.

\end{eabstract}
\begin{jabstract}
逆コンパイラはコンパイル後のバイナリデータからソースコードを復元するためのツールである。
様々な逆コンパイラが存在するが、既存の逆コンパイラはしばしば人間にとって分かりにくいソースコードを出力する。
そこで本論文では、統計的機械翻訳の技術を逆コンパイラに用いた、解析者にとってわかりやすいソースコードを出力する逆コンパイラを提案する。
具体的には、機械翻訳において有用とされる注意機構付き再帰型ニューラルネットワークを用いる。
実験では、オープンソースプロジェクトから収集したソースコードとそのバイナリデータを用いてニューラルネットワークを学習し、その性能を検証した。
実験の結果、注意機構を用いることににより逆コンパイル結果における若干の精度向上が認められたが、実用的な精度までは至らなかった。
\end{jabstract}

@@

\begin{figure}
	\begin{tabular}{cc}
		\begin{minipage}[c]{0.5\hsize}
		\fbox{
			\lstinputlisting[breaklines=false]{insert_list.c}
			}
			\caption*{Sample C source code}
		\end{minipage}
		\\ \\
		\begin{minipage}[b]{0.5\hsize} %title={Result of the Hopper Decompiler}
			\begin{framed}
			\lstinputlisting[]{snowman.c}
			\end{framed}
			\caption*{Result of the Snowman Decompiler}
		\end{minipage}
		\begin{minipage}[b]{0.5\hsize}
			\begin{framed}
			\lstinputlisting[breaklines=false]{recstudio.c}
			\end{framed}
			\caption*{Result of the REC Studio Decompiler}
		\end{minipage}
	\end{tabular}
	% \vspace*{-0.6cm}
	\caption{Decompilation results of existing decompilers for C source code}
	\label{fig:cw}
\end{figure}
Figure~\ref{fig:cw} shows an example of decompilation results of the Snowman decompiler version v0.1.3 and the REC Studio decompiler version 4.1 
for a sample C language code. The sample C code is compiled by the GNU C Compiler version 4.8.5, with the optimization level \texttt{-O0}.
The REC decompiler missed the structure of the variable \texttt{root}.
It also failed to estimate the number of arguments and the return value of the function.
The Snowman Decompiler succeeded in reconstructing the structure of the variable as \texttt{struct s0}.
It also succeeded in estimating the number of arguments and the return value as semantically correct expressions. 
However the statement \texttt{return 0;} is reconstructed to a complex combination of statements.

>>>>

% documentclass[senior,final,11pt]{iscs-thesis}
\documentclass[11pt]{jarticle}
%論文の種類とフォントサイズをオプションに
%\usepackage{graphicx}% 必要に応じて
%\usepackage{mysettings}% 自分用設定

@@

%-------------------
% \title{Neural Network based Decompiler and its Evaluation}
\title{ニューラルネットを用いた逆コンパイラとその評価}

% __NAMES__

@@

\usepackage{listings}
\usepackage{cite}
\usepackage{url}
\usepackage[numbers]{natbib}
\usepackage[dvipdfmx]{graphicx}
\usepackage{framed}
\usepackage{caption}
\usepackage{comment}

@@

\begin{document}
\begin{comment}
\begin{eabstract}
A decompiler is a tool for recovering a source code from compiled binary data.
There are various decompilers, but they often output a source code that is not intelligible to humans. 
In this thesis, we apply machine translation techniques to generate human intelligible decompiled source codes. 
More specifically, we propose to use recurrent neural networks with attention, 
which have been demonstrated to be useful for machine translation. 
In experiments, we use source codes collected from open source projects and their binary data for training and evaluating the neural networks.
From the experiment, we can observe little improvement in the result of decompilation with the use of attention.
Unfortunately, the results were not still good enough for practical use.

\end{eabstract}
\end{comment}
\begin{abstract}
逆コンパイラはコンパイル後のバイナリデータからソースコードを復元するためのツールである。
様々な逆コンパイラが存在するが、既存の逆コンパイラはしばしば人間にとって分かりにくいソースコードを出力する。
そこで本論文では、統計的機械翻訳の技術を逆コンパイラに用いた、解析者にとってわかりやすいソースコードを出力する逆コンパイラを提案する。
具体的には、機械翻訳において有用とされる注意機構付き再帰型ニューラルネットワークを用いる。
実験では、オープンソースプロジェクトから収集したソースコードとそのバイナリデータを用いてニューラルネットワークを学習し、その性能を検証した。
実験の結果、注意機構を用いることににより逆コンパイル結果における若干の精度向上が認められたが、実用的な精度までは至らなかった。
\end{abstract}

@@

\begin{figure}
	\begin{tabular}{cc}
		\begin{minipage}[c]{0.5\hsize}
		\fbox{
			% \lstinputlisting[breaklines=false]{insert_list.c}
			}
			\caption*{Sample C source code}
		\end{minipage}
		\\ \\
		\begin{minipage}[b]{0.5\hsize} %title={Result of the Hopper Decompiler}
			\begin{framed}
			\lstinputlisting[]{snowman.c}
			\end{framed}
			\caption*{Result of the Snowman Decompiler}
		\end{minipage}
		\begin{minipage}[b]{0.5\hsize}
			\begin{framed}
			% \lstinputlisting[breaklines=false]{recstudio.c}
			\end{framed}
			\caption*{Result of the REC Studio Decompiler}
		\end{minipage}
	\end{tabular}
	% \vspace*{-0.6cm}
	\caption{Decompilation results of existing decompilers for C source code}
	\label{fig:cw}
\end{figure}
Figure~\ref{fig:cw} shows an example of decompilation results of the Snowman decompiler version v0.1.3 and the REC Studio decompiler version 4.1 
for a sample C language code. The sample C code is compiled by the GNU C Compiler version 4.8.5, with the optimization level \texttt{-O0}.
The REC decompiler missed the structure of the variable \texttt{root}.
It also failed to estimate the number of arguments and the return value of the function.
The Snowman Decompiler succeeded in reconstructing the structure of the variable as \texttt{struct s0}.
It also succeeded in estimating the number of arguments and the return value as semantically correct expressions. 
However the statement \texttt{return 0;} is reconstructed to a complex combination of statements.
